---
---
---

#### Useful constants

```{r}
set.seed(022)
library(markovchain)
TLN <- 022
true <- TRUE
```

# Problem 1

### Estimate numerically the probability of the event that your TLN occurs in a random digit sequence $d_1d_2d_3 . . . d_n$.

```{r}
how_often <- function(size = 1e7) {
  res <- 0
  numbers <- sample(0:9, size, replace = true)
  for (i in 1:(size - 2)) {
    if (numbers[i] == 0 &&
      numbers[i + 1] == 2 &&
      numbers[i + 2] == 2) {
      res <- res + 1
    }
  }
  return(res / size)
}



simulate <- function(size) {
  res <- 0
  numbers <- sample(0:9, size, replace = true)
  for (i in 1:(size - 2)) {
    if (numbers[i] == 0 &&
      numbers[i + 1] == 2 &&
      numbers[i + 2] == 2) {
      return(1)
    }
  }
  return(0)
}

n_values <- c(100, 200, 1000)
N <- 1e3

compute_estimate <- function(n, times = 1) {
  res <- seq(1, times, by = 1)
  for (i in 1:times) {
    res[i] <- simulate(n)
  }
  return(res)
}


emp_lim_small <- compute_estimate(n_values[1], times = N)
emp_lim_medium <- compute_estimate(n_values[2], times = N)
emp_lim_large <- compute_estimate(n_values[3], times = N)
mean(emp_lim_small)
mean(emp_lim_medium)
mean(emp_lim_large)
```

### Limiting probability

```{r}
transition_matrix <- matrix(data = c(0.9, 0.1, 0, 0, 0.8, 0.1, 0.1, 0, 0.8, 0.1, 0, 0.1, 0.9, 0.1, 0, 0), nrow = 4, ncol = 4, byrow = true)
transition_matrix

mc <- new("markovchain", transitionMatrix = transition_matrix)
print("Real matrix and limiting probabilities: ")
steadyStates(mc)

print("Experimantal limiting probability:")
mean(how_often(1e7))
```

#### Results are very close!

### Determining the sample size

$$
\Phi(- \frac{\sqrt{n} \cdot \epsilon}{\sigma}) = \frac{\alpha}{2}
$$

$$
n = (-\frac{\Phi^{-1}(\alpha/2) \cdot \sigma} {\epsilon})^2
$$

```{r}
check_conf_level <- function(n, enough, theory) {
  times <- 500
  results <- seq(1, times, by = 1)
  for (i in 1:times) {
    checking <- seq(1, enough, by = 1)
    for (ind in 1:enough) {
      checking[ind] <- compute_estimate(n)
    }
    results[i] <- mean(checking)
  }
  return(length(results[abs(results - theory) < 0.03]) / times)
}

for (n_ind in 1:3) {
  n <- n_values[n_ind]
  theretical_sample <- compute_estimate(n, 1e5)
  theoretical_mean <- mean(theretical_sample)
  theoretical_sd <- sd(theretical_sample)
  enough <- (theoretical_sd * qnorm(0.025) / 0.03)^2
  cat("For n value", n, "the required sample size is ", enough, "\n")
  res <- check_conf_level(n, enough, theoretical_mean)
  cat("which gives ", res, " confidence level", "\n")
}

```

# Problem 2

### Make necessary amendments to the transition probabilities matrix P above and solve the corresponding system to find the expected time $E(T)$ till absorption

```{r}
transition_matrix <- matrix(data = c(0.9, 0.1, 0, 0, 0.8, 0.1, 0.1, 0, 0.8, 0.1, 0, 0.1, 0, 0, 0, 1), nrow = 4, ncol = 4, byrow = true)
transition_matrix


mc <- new("markovchain", transitionMatrix = transition_matrix)
meanAbsorptionTime(mc)
```

Here we see that the expected absorption time is 1000.

### Estimate numerically the expected length E(T) till the first occurrence of the TLN by running a sufficiently large number N of experiments.

```{r}
until_first_occurrence <- function() {
  size <- 1e3
  numbers <- sample(0:9, size, replace = true)
  for (i in 1:(size - 2)) {
    if (numbers[i] == 0 &&
      numbers[i + 1] == 2 &&
      numbers[i + 2] == 2)
    {
      return(i + 2)
    }
  }
  return(size + until_first_occurrence())
}

generate_alot_first <- function(size) {
  just_sample <- seq(1, size, by = 1)
  for (i in 1:size) {
    just_sample[i] <- until_first_occurrence()
  }
  return(just_sample)
}


N <- 2 * 1e4
mean(generate_alot_first(N))
```

We see that results all the time are very close.

### Find the sample size N which guarantees that the absolute error \|θˆ−θ\| of the estimate does not exceed 10 with confidence level of at least 95 percent.

$$
P(|\hat{\theta} - \theta| ≥ c) ≤ \frac{\sigma_x^2}{c^2}
$$

$$
P(|\hat{\theta} - \theta| ≤ c) ≥  1 -\frac{\sigma_x^2}{c^2}
$$

$$
\theta = 1000; \space c = 10 
$$

The only missing value is $\sigma$ --- standard deviation of $\hat{\theta}$. The dependence between standard error and standard deviation is $\sqrt{n}$. If we have the sample of size $n$, then $\sigma_x = \frac{\sigma}{\sqrt{n}}, \sigma_x$ - standard deviation of the sample.

$$
1 - \frac{\sigma_x}{c^2} = 0.95 => \sigma_x = c\sqrt{0.05} \approx 2.236
$$

$$
\sqrt{n} = \frac{\sigma}{\sigma_x}
$$

Seeking theoretical sigma:

```{r}
number_of_simulations <-  1*1e2

sample_size <- 0.5 * 1e3

values <- c(1, number_of_simulations, by=1)

for (i in 1:number_of_simulations){
  tmp <- 0
  for(j in 1:sample_size){
    tmp <- tmp + until_first_occurrence()
  }
  values[i] <- tmp / sample_size
}

```

```{r}
sd(values) * sqrt(sample_size)
```

$$
\sqrt{n} = \frac{1000}{2.236} \approx 447; n = 199809 \approx 2 * 10^5
$$

This is too large of a value because Chebyshev does not give close results and overestimates the answer. The real value is about 3 times less:

```{r}
enough <- 3.8*1e4

times <- 40
res <- 0
for(i in 1:times){
  tmp <- seq(1, enough, by=1)
  for (ind in 1:enough) {
    tmp[ind] <- until_first_occurrence()
  }
  if (abs(mean(tmp) - 1000) < 10){
    res <- res + 1
  }
}
res / times
```
